\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}

\title{Data Sources for Project: Analyzing Sports Performance and Social Media Trends}
\author{Ebenezer Olubayode}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}
This document outlines the data sources of interest for a comprehensive project aimed at generating insights into team and player performances, as well as analyzing social media trends related to the NBA. The goal is to utilize advanced data scraping techniques to gather relevant information from both structured and unstructured data sources.

\section{Data Sources}

\subsection{FBref for Sports Analytics}
FBref.com is a premier source for football (soccer) statistics and data. For the purpose of analyzing team and player performances, data from FBref will be scraped. This data includes but is not limited to player statistics, team performance metrics, game outcomes, and historical data archives. The insights derived from this analysis could significantly contribute to understanding the dynamics of player contributions and team strategies.

\subsection{Twitter Data for Social Media Analytics}
Social media platforms, particularly Twitter, are rich sources of public sentiment and topical discussions. By scraping Twitter data for mentions related to the NBA, including specific hashtags and keywords, we aim to perform social media analytics using topic modeling techniques. This analysis will help identify the most widely discussed topics, public sentiment towards players and teams, and the overall impact of events on social media discourse.

\section{Objective}
The primary objective of this project is to leverage data scraping methodologies to access and analyze specific datasets from FBref and Twitter. Through this, we aim to gain deeper insights into:

\begin{itemize}
    \item Team and player performances, leveraging statistical data from FBref.
    \item Social media trends and public sentiment related to the NBA, utilizing data from Twitter.
\end{itemize}

\section{Conclusion}
Accessing and analyzing data from FBref and Twitter will provide a multi-dimensional view of sports analytics and social media trends. This project not only aims to uncover insights into sports performances but also seeks to understand the public discourse surrounding sports events, especially in the context of the NBA.

\end{document}
